### 2013-10-11 Disseration/article.tex
added figures, class diagrams

\documentclass[a4paper,12pt]{article}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage[top=2.5cm, bottom=2.5cm, left=2.5cm, right=2.5cm]{geometry}
\usepackage{cite}

%\setlength{\parskip}{6pt}%
\linespread{1.5}

\newcommand{\superscript}[1]{\ensuremath{^{\textrm{#1}}}}
\newcommand{\subscript}[1]{\ensuremath{_{\textrm{#1}}}}

%opening
\title{Object Categorization based on RGBD data}
\author{Adam Kosiorek}
\bibliographystyle{abbrv}
\begin{document}

\maketitle

\include{streszczenie}
\include{abstract}
\include{introduction} %aka motivation
\include{bow}
\include{libs}
\include{datasets}
\include{architecture}
\include{setup}
\include{experiments}
\include{conclusion}

\bibliography{article}

\end{document}