### 2013-10-12 Disseration/abstract.tex
pfhrgb, changed layout of the diss

\begin{center}
\begin{LARGE}\textbf{Abstract}\end{LARGE}
\end{center}

\vspace{1.0cm}

Properties of special generalized functions and short-range selectors which are indispensable in the Shina Tan description of interacting Fermi systems [Annals of Physics 323, 2952 (2008)] are analyzed in details. Also all details of the proof of the Shina Tan energy theorem and properties of the contact are presented. The manuscript should be interesting for researchers who want to understand all elements of the Shina Tan extraordinary work.