### 2013-10-12 Disseration/article.tex

\documentclass[12pt]{report}
\usepackage[utf8]{inputenc}
%\usepackage{polski}
\usepackage[T1]{fontenc}
\usepackage[a4paper, top=2.5cm, bottom=2.5cm, left=2.5cm, right=2.5cm]{geometry}
\usepackage{cite}
\usepackage{graphicx}
%\usepackage[polish]{babel}
\usepackage{color}

%\setlength{\parskip}{6pt}%
\linespread{1.5}

\newcommand{\superscript}[1]{\ensuremath{^{\textrm{#1}}}}
\newcommand{\subscript}[1]{\ensuremath{_{\textrm{#1}}}}

\bibliographystyle{abbrv}
\begin{document}
\pagestyle{empty}
\color{white} . \color{black}
\vspace{0.2cm}

\begin{center}
\begin{Huge}Warsaw University of Technology
\end{Huge}
\end{center}

\begin{center}
\begin{LARGE}Faculty of Mechatronics

\vspace{0.5cm}
Institute of Automation and Robotics
\vspace{0.5cm}
\end{LARGE}
\end{center}
\begin{center}\includegraphics[width=5cm,angle=0]{figs/logopw}\end{center}

\vspace{1.0cm}

\begin{center}
\begin{LARGE}\textbf{Adam Kosiorek}\end{LARGE}
\end{center}

\vspace{0.5cm}

\begin{center}
\begin{Huge}\textbf{3D Object Classification based on RGBD images}
\vspace{0.5cm}
\end{Huge}
\end{center}

\vspace{0.5cm}

\begin{center}
\begin{Large}bachelor thesis written under supervision of\end{Large}

\vspace{0.2cm}

\begin{LARGE}prof. dr. hab. Barbara Siemiątkowska\end{LARGE}
\end{center}

\vspace{1.8cm}

\begin{center}
\begin{Large}Warsaw 2014\end{Large}
\end{center}

\pagebreak

\color{white} . \color{black}
\vspace{0.2cm}
\vspace{1.5cm}
\pagebreak
\pagestyle{plain}

\include{streszczenie}
\include{abstract}

\tableofcontents

\chapter{Introduction}
\addcontentsline{toc}{chapter}{Introduction}
\include{introduction} %aka motivation
\include{bow}
\include{libs}
\include{datasets}

\chapter{Design}
\addcontentsline{toc}{chapter}{Design}
\include{architecture}
\include{setup}

\chapter{Experiments and results}
\addcontentsline{toc}{chapter}{Experiments and results}
\include{experiments}
\include{conclusion}

\listoffigures
\listoftables

\cite{OpenCV}

\bibliography{article}

\end{document}