### 2013-10-12 Disseration/experiments.tex

\section{Experiments}

	The Bag of Words classification pipeline consists of four previously discussed steps. For there are many different algorithms suitable for each part of the pipeline, an extensive research is required in order to discover the best possible combination in a particular setting. The number of options grows combinatorial with number of candidates for each processing step. Moreover, many algorithms have different parameters that should be adjusted for the given conditions. Evaluation of every possibility is unfeasible due to the lack of time and computational power required. Therefore it was chosen to evaluate algorithms with their default parametrs under fixed conditions. Only the winning methods qualify for the parameter tuning.

\subsection{Detectors}

	ISS3D is generally better than SIFT

\subsection{Descriptors}

	FPFH is faster than both PFH and PFHRGB. PFH does not give any advantages over FPFH. PFHRGB delievers considerably higher accuracy.

\subsection{Codebook}

Only kmeans tested with varying numbers of clusters i.e. vocabulary size

\subsection{B3DO}

\subsection{tokyo}