### 2014-01-06 Disseration/article.tex

%TODO
%	- run cluster and descriptors tests
%	- find out what's wrong with sift detector
%	- run detector tests
%	- prepare strategy pattern class diagram and describe it
%	- write conclusions
%	- describe libraries
%	- write abstracts
%	- go through introduction
%
%
%

\documentclass[12pt]{report}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage[a4paper, top=2.5cm, bottom=2.5cm, left=2.5cm, right=2.5cm]{geometry}
\usepackage{cite}
\usepackage{graphicx}
\usepackage{color}
\usepackage[font=footnotesize]{caption}
\usepackage{amsmath}
\usepackage{hyperref}
\usepackage{paralist}

\linespread{1.5}

\newcommand{\superscript}[1]{\ensuremath{^{\textrm{#1}}}}
\newcommand{\subscript}[1]{\ensuremath{_{\textrm{#1}}}}
\newcommand\rednote[1]{\textcolor{red}{#1}}

\makeatletter
% Redefine the \chapter* header macro to remove vertical space
\def\@makeschapterhead#1{%
  %\vspace*{50\p@}% Remove the vertical space
  {\parindent \z@ \raggedright
    \normalfont
    \interlinepenalty\@M
    \Huge \bfseries  #1\par\nobreak
    \vskip 20\p@
  }}
\makeatother

\bibliographystyle{abbrv}
\begin{document}

\include{titlepage}
\include{zyciorys}

\pagestyle{plain}

\include{abstract}

\tableofcontents
\include{introduction}
\include{architecture}
\include{experiments}

\listoffigures
\listoftables

\bibliography{article}

\end{document}