### 2014-01-06 Disseration/titlepage.tex

\pagestyle{empty}
\color{white} . \color{black}
\vspace{0.2cm}

\begin{center}
\begin{Huge}Warsaw University of Technology
\end{Huge}
\end{center}

\begin{center}
\begin{LARGE}Faculty of Mechatronics

\vspace{0.5cm}
Institute of Automation and Robotics
\vspace{1.0cm}
\end{LARGE}
\end{center}
\begin{center}\includegraphics[width=5cm,angle=0]{figs/logopw}\end{center}

\vspace{0.5cm}

\begin{center}
\begin{LARGE}\textbf{Adam Kosiorek}\end{LARGE}
\end{center}

\vspace{0.5cm}

\begin{center}
\begin{Huge}\textbf{3D Object Classification based on RGBD images}
\vspace{0.5cm}
\end{Huge}
\end{center}

\vspace{0.5cm}

\begin{center}
\begin{Large}bachelor thesis written under supervision of\end{Large}

\vspace{0.2cm}

\begin{Large}prof. dr. hab. Barbara Siemiątkowska\end{Large}
\end{center}

\vspace{1.8cm}

\begin{center}
\begin{Large}Warsaw 2014\end{Large}
\end{center}