### 2014-02-04 Disseration/zyciorys.tex

    \vspace*{1\baselineskip}
    \begin{tabular}{p{5cm} p{10cm}}
    \begin{minipage}{5cm}
    \center
    \includegraphics[width=4.5cm]{figs/ja}
    \end{minipage}
    &
    \begin{minipage}{12cm}
\par\noindent\vspace{1\baselineskip}
    \begin{flushleft}
    \par\noindent\vspace{1\baselineskip}
    \begin{tabular}[h]{l l}
    {\normalsize\it Degree:} & Automatic Control Engineering and Robotics
    \end{tabular}
    \par\noindent\vspace{1\baselineskip}
    \begin{tabular}[h]{l l}
    {\normalsize\it Major:} & Robotics
    \end{tabular}
    \par\noindent\vspace{1\baselineskip}
    \begin{tabular}[h]{l l}
    {\normalsize\it Birth date:} & {\normalsize 28 February 1991}
    \end{tabular}
    \par\noindent\vspace{1\baselineskip}
    \begin{tabular}[h]{l l}
    {\normalsize\it Study starting date:} & {\normalsize 1 October 2010 r.}
    \end{tabular}
    \par\noindent\vspace{1\baselineskip}
    \end{flushleft}
    \end{minipage}
    \end{tabular}
    \vspace*{1\baselineskip}
    \begin{center}
	{\large\bfseries Biography}\par\bigskip
    \end{center}

	I was born on 28 February 1991 in Olsztyn, Poland. I attended 5\superscript{th} Secondary School by the name of ``The Common Europe'' in Olsztyn. I started a Bachelor of Engineering degree on the Faculty of Mechatronics of Warsaw University of Technology on 1 October 2010. I did an internship in Faurecia~R\&D~S.A. in 2012. I joined IBM Poland Sp.~z~o.~o. in January, 2013. Then, I started working full-time in Samsung R\&D Centre Poland as a Computer Vision Engineer. I am an avid long distance runner and a keen reader of fantastic literature and thrillers.

    \par
    \vspace{3\baselineskip}
    \hfill\parbox{15em}{{\small\dotfill}\\[-.3ex]
    \centerline{\footnotesize signature}}\par
    \vspace{3\baselineskip}